---
title: "p8105_hw6_az2852"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(stringr)
library(broom)
```

```{r}
set.seed(1)
```

# Problem 2
Load and prepare the homicide data.
```{r}
homicide = read_csv("data/homicide-data.csv")

# Data preparation
homicide_df = homicide %>%
  mutate(city_state = str_c(city, state, sep = ", ")) %>%
  mutate(solved = ifelse(disposition == "Closed by arrest", 1, 0)) %>%
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))
  ) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  mutate(victim_age = as.numeric(victim_age))
```
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
baltimore_data = homicide_df %>%
  filter(city_state == "Baltimore, MD")

# save the result as an R object
baltimore_model = glm(
  solved ~ victim_age + victim_sex + victim_race,
  data = baltimore_data,
  family = binomial(link = "logit")
)
# apply broom::tidy
baltimore_results = broom::tidy(baltimore_model)

# Calculate adjusted odds ratio (OR) and confidence intervals for male vs female victims
baltimore_or = baltimore_results %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    odds_ratio = exp(estimate), # Adjusted OR
    conf_low = exp(estimate - 1.96 * std.error), # Lower CI
    conf_high = exp(estimate + 1.96 * std.error) # Upper CI
  ) %>%
  select(term, odds_ratio, conf_low, conf_high)

baltimore_or
```

Regression model for each city.
```{r}
city_results = homicide_df %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, \(x) glm(solved ~ victim_age + victim_sex + victim_race, 
                            data = x, 
                            family = binomial(link = "logit"))), 
    tidy_model = map(model, broom::tidy) 
  ) %>%
  unnest(tidy_model) %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(
    odds_ratio = exp(estimate), 
    conf_low = exp(estimate - 1.96 * std.error), 
    conf_high = exp(estimate + 1.96 * std.error) 
  ) %>%
  select(city_state, odds_ratio, conf_low, conf_high)

city_results
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r fig.height= 8, fig.width=8}

# Sort cities by OR
results = city_results %>%
  arrange(odds_ratio) %>%
  mutate(city_state = fct_reorder(city_state, odds_ratio))

# Plot the ORs and CIs
ggplot(results, aes(x = odds_ratio, y = fct_reorder(city_state, odds_ratio))) +
  geom_point(color = "blue", size = 2) +                    
  geom_errorbarh(aes(xmin = conf_low, xmax = conf_high),       
                 height = 0.2, color = "darkgray") +
  scale_x_log10() +                                         
  labs(
    title = "Estimated Odds Ratios for Solving Homicides by City",
    x = "Odds Ratio (Log Scale)",
    y = "City",
    caption = "ORs and CIs for resolving homicides comparing male victims to female victims"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    plot.title = element_text(size = 14, face = "bold"),
    plot.caption = element_text(size = 10)
  )

```

# Problem 3
load and clean the data
```{r}
birthweight = read_csv("data/birthweight.csv")
```
```{r}
birthweight_df = birthweight %>%
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("Male", "Female")),
    malform = factor(malform, levels = c(0, 1), labels = c("Absent", "Present")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown"))
  )
```

check for missing values
```{r}
summary(birthweight_df)
```
Fit regression models of birth weight. We want to explore the relationship between different biological and maternal health factors and baby weight. To do so, we begin with predictors gestational age, smoking status, mother's delivery weight, baby's head circumference, length, and sex. We found a main effect for gestational age, mother's delivery weight, baby's head circumference, length, and sex respectively, and significant interactions between head circumference, length, and sex. We did not find a main effect for smoking nor any interaction effect of smoking with other variables. 
```{r}
# Regression model predicting birth weight based on gestational age, delivery weight, baby's head circumference, length, and sex and the interaction between those three.
model_1 = lm(bwt ~ gaweeks + delwt + bhead * blength * babysex, data = birthweight_df)

summary(model_1)

# Add predictions and residuals
data_with_preds = birthweight_df %>%
  add_predictions(model_1, var = "fitted_bwt") %>%
  add_residuals(model_1, var = "residuals_bwt")

# Plot residuals against fitted values
ggplot(data_with_preds, aes(x = fitted_bwt, y = residuals_bwt)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(
    title = "Residuals vs Fitted Values for Birth Weight Model",
    x = "Fitted Values (Predicted Birth Weight)",
    y = "Residuals"
  ) +
  theme_minimal()

# Check residual normality with a histogram & Q-Q plot
ggplot(data_with_preds, aes(x = residuals_bwt)) +
  geom_histogram(binwidth = 50, fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Distribution of Residuals",
    x = "Residuals",
    y = "Frequency"
  )

qqnorm(data_with_preds$residuals_bwt)
qqline(data_with_preds$residuals_bwt, col = "red")
```

The residual plot show a random pattern with points evenly distributed above and below the horizontal line. There are a few outliers with high residuals.

The residual histogram appears to be roughly bell-shaped but with outliers on the higher end.

The QQ plot show that most points fall perfectly on the diagonal, except a slight deviation at the tails, but this may not be a concern considering the size of our data (4342 observations).

## compare the proposed model with 2 alternative models
```{r}
# 2. Alternative model 1: length at birth and gestational age (main effects only)
model_2 = lm(bwt ~ blength + gaweeks, data = birthweight_df)

summary(model_2)
# 3. Alternative model 2: head circumference, length, sex, and all interactions
model_3 = lm(bwt ~ bhead * blength * babysex, data = birthweight_df)

summary(model_3)
```
Now, do cross-validation between 3 models
```{r}
# create training and testing samples
cv_df = 
  crossv_mc(birthweight_df, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df %>% pull(train) %>% nth(1) %>% as.tibble()
```
```{r}
cv_res = 
  cv_df |> #fit the regression models on training data
  mutate(
    model_1  = map(train, \(df) lm(bwt ~ gaweeks + delwt + bhead * blength * babysex, data = df)),
    model_2  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    model_3  = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))) |> #calculate the RMSE for the model fit on test data
  mutate(
    rmse_1 = map2_dbl(model_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_2 = map2_dbl(model_2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_3 = map2_dbl(model_3, test, \(mod, df) rmse(model = mod, data = df)))

cv_res
```
see the distribution of RMSE
```{r}
cv_res |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model))|> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Model 1 seems to perform slightly better than model 3.
